---
title: "data-storyteller"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, results ='show',include=TRUE,messages=FALSE)

library(tidyverse)
library(forcats)
library(ggplot2)
library(dplyr)
library(stargazer)
library(caret)
library(modelr)
```

# Introduction

The topic of this research is to focus on understanding how students are performing from a collected dataset. By performing exploratory data analysis. The researcher would have the ability to analyze and understand the performance statistically. 

Students' performance is critical in many ways. To school and teachers, understanding their performance will help the instructors adjust and pace of teaching and related course materials. To parents, understanding students' performance enables them to keep track of how their kids are doing in school. The performance of students is also important to the public. The public wants to understand what schools would have better infrastructure in supporting students' learning progress and a better chance of improving their performance. 

# Analysis and Models
## Reading in the data.

The sample data is a rather small dataset. The data contains the progress of students taking the same math course this semester. There are five schools (A, B, C, D, and E). The semester is about ¾ of the way through. 
There are 8 variables in the dataset. The variable names and sample data are presented below. The variable school represents where the student took the course. The variable section represents which section of the student was in. Variable very ahead, middling, behind, more behind, very behind, and completed represent how students are performing.

```{r getData, include=TRUE}
#replacing any possible blank entries as 'NA'.

storyteller<- read_csv("data-storyteller.csv", na = c(""))

#checking data types to see what may need changing
str(storyteller)
```


# Data Cleaning
The performance variables were recorded in below fashion.
Note: The School column is of the character type and it should be a factor. Other notes below in comments. 

```{r cleanData1, include=TRUE}
storyteller$School<-factor(storyteller$School)

#The section column is of the Numeric type and should be be a factor instead

storyteller$Section<-factor(storyteller$Section)

#Each of the remaining columns is a discrete count of the students in each category.
#As it is not continuous the columns 'Very Ahead', 'Middling', 'Behind', 'More behind', 'Very behind'
#and 'Completed' should all be integers.

storyteller$`Very Ahead +5`<-as.integer(storyteller$`Very Ahead +5`)
storyteller$`Middling +0`<-as.integer(storyteller$`Middling +0`)
storyteller$`Behind -1-5`<-as.integer(storyteller$`Behind -1-5`)
storyteller$`More Behind -6-10`<-as.integer(storyteller$`More Behind -6-10`)
storyteller$`Very Behind -11`<-as.integer(storyteller$`Very Behind -11`)
storyteller$Completed<-as.integer(storyteller$Completed)

```

In the data cleaning section, it's critical to understand how the researcher identifies the data types of each variable. The researcher defines variable school and section to be a factor. In other words, those two variables are categorical variables. Categorical variables help us label the data. Other variables such as very ahead, middling, behind, more behind, very behind, and completed represent the discrete count of the students in each category. The data should all be integers.

One of the important processes is to identify the null value or missing data. By running the command, we can see there isn’t any missing data. The dataset is very clean.

## Organizing the Data Structure

```{r Organize}

#Reordering columns to get a cleaner picture. I.E. 'Completed' being shifted to the other side.
#And section being a unique identifier is moved to the leftmost column.

storytellerTemp<-storyteller[,c(2,1,8,3,4,5,6,7)]

storyteller<-storytellerTemp

#displaying top 5 rows
head(storyteller)

```

## Missing Data

```{r cleanData2, include=TRUE}

#Checking for any NA values
sum(is.na(storyteller))

#There are no NA values in this dataset.

#The dataset is cleaned
head(storyteller)

```

# EDA and Data Viz

The research of this dataset will be mainly focused on descriptive analysis. The researcher will compose many plots and graphs to understand the data descriptively.

This bar chart represents the number of sections for each school. By reading the plot, we understand A has the most sections. The ranking is A, B, C, D, and E. E has the least sections.

```{r viz, include=TRUE}

# This bar chart represents the number of sections for each school. By reading the plot, we understand A has the most sections. The ranking is A, B, C, D, and E. E has the least sections.

SchoolValues<-c(length(which(storyteller$School=='A')), length(which(storyteller$School=='B')), length(which(storyteller$School=='C')), length(which(storyteller$School=='D')), length(which(storyteller$School=='E')))

barplot(SchoolValues, names.arg = c('A', 'B', 'C', 'D', 'E'), main='Number of sections Per School')

#This bar plot represents the numbers of students who completed the courses on time and which section they are from. The researcher can’t identify any types of distribution in this graph. However, section 30 seems to have more students who completed the courses on time followed by 10, 27, and 7.

barplot(storyteller$Completed, main='#completed students / section', names.arg = c(1:30))
summary(storyteller$Completed)
```

## Further Data Cleaning based on EDA and Viz

```{r cleanData3, include=TRUE}

#Plotting section and Very Ahead and summarizing
#This graph presents the number of students who are very ahead comparing to other students. The plot shows there aren't any students who are very ahead in the course.

barplot(storyteller$`Very Ahead +5`, main='number of very ahead students per section', names.arg = c(1:30))

summary(storyteller$`Very Ahead +5`)
#plotting section and Middling and summarizing
#This bar plot represents how many students are in the middle status. Section 7 takes the lead followed by 4 and 10.

barplot(storyteller$`Middling +0`, main='number of Middling students per section', names.arg = c(1:30))

summary(storyteller$`Middling +0`)
#plotting section and Behind and summarizing
#This plot represents the number of students who are behind. Section 30 has a number of students who are behind.

barplot(storyteller$`Behind -1-5`, main='number of Behind students per section', names.arg = c(1:30))

summary(storyteller$`Behind -1-5`)
#plotting More Behind and Very Behind by section
#Below two plots represented how many students are more and very behind in the course. Section 2 to section 14 has a higher number of students who are more behind and very behind.

barplot(storyteller$`More Behind -6-10`, main='number of More Behind students per section', names.arg = c(1:30))


summary(storyteller$`More Behind -6-10`)
#Plotting section 

barplot(storyteller$`Very Behind -11`, main='number of Very Behind students per section', names.arg = c(1:30))

summary(storyteller$`Very Behind -11`)
#determining the number of students in each category
#This plot shows that most of the students are behind in the course. Also, the statistics show that 20% of students completed the course, 0% of students are very ahead, 14% of students are middling, 47% of students are behind, 6% of students are more behind, and 13% of students are very behind. That being said, there are almost 70% of students who are behind in the course.
StudentSums<-colSums(storyteller[,3:8])

sum(StudentSums)

#determining the amount of students in each section

SectionSums<-rowSums(storyteller[,3:8])

data.frame(SectionSums)

#Creating a barplot to show distribution
#This plot shows that most of the students are behind in the course

StudentSums<-colSums(storyteller[,3:8])
barplot(StudentSums, main="Student totals across all categories")

```

Now, we would think the course must be very challenging that most of the students are behind. Is that true? How are the students performing in each school? The below plots show the counts of students' progress for each school.

## EDA (cont.)

```{r EDA, include=TRUE}

StudentSums/sum(StudentSums)

```

The statistics show that 20% of students completed the course, 0% of students are very ahead, 14% of students are middling, 47% of students are behind, 6% of students are more behind, and 13% of students are very behind. That being said, there are almost 70% of students who are behind in the course.

```{r middling, include=TRUE}

plot(storyteller$Completed, storyteller$`Middling +0`)

```

```{r Behind, include=TRUE}

plot(storyteller$Completed, storyteller$`Behind -1-5`)

```

```{r moreBehind, include=TRUE}

plot(storyteller$Completed, storyteller$`More Behind -6-10`)

```

```{r veryBehind, include=TRUE}

plot(storyteller$Completed, storyteller$`Very Behind -11`)

```

```{r corr, include=TRUE}

cor(storyteller[3:7], use="complete.obs")

```

```{r boxplot, include=TRUE}

boxplot(storyteller$Completed, main="Completed")
boxplot(storyteller$`Middling +0`, main="Middling")
boxplot(storyteller$`Behind -1-5`, main="Behind")
boxplot(storyteller$`More Behind -6-10`, main="More Behind")
boxplot(storyteller$`Very Behind -11`, main="Very Behind")

```

```{r log, include=TRUE}

plot(storyteller$Completed, log(storyteller$Completed), main="Completed")
plot(storyteller$`Middling +0`, log(storyteller$`Middling +0`), main="Middling")
plot(storyteller$`Behind -1-5`, log(storyteller$`Behind -1-5`), main="Behind")
plot(storyteller$`More Behind -6-10`, log(storyteller$`More Behind -6-10`), main="More Behind")
plot(storyteller$`Very Behind -11`, log(storyteller$`Very Behind -11`), main="Very Behind")

```

```{r orderCat, include=TRUE}

sortedCompleted <- storyteller[order(-storyteller$Completed),] 
sortedCompleted

sortedMoreBehind <- storyteller[order(-storyteller$`More Behind -6-10`),] 
sortedMoreBehind

```

Now, we would think the course must be very challenging that most of the students are behind. Is that true? How are the students performing in each school? The below plots show the counts of students' progress for each school.

```{r groupAnalysis, include=TRUE}

storytellerA<-storyteller[which(storyteller$School == "A"),]
storytellerB<-storyteller[which(storyteller$School == "B"),]
storytellerC<-storyteller[which(storyteller$School == "C"),]
storytellerD<-storyteller[which(storyteller$School == "D"),]
storytellerE<-storyteller[which(storyteller$School == "E"),]

StudentSumsA<-colSums(storytellerA[3:8])
StudentSumsA

barplot(StudentSumsA, main = "School A")

StudentSumsB<-colSums(storytellerB[3:8])
StudentSumsB

barplot(StudentSumsB, main = "School B")

StudentSumsC<-colSums(storytellerC[3:8])
StudentSumsC

barplot(StudentSumsC, main = "School C")

StudentSumsD<-colSums(storytellerD[3:8])
StudentSumsD

barplot(StudentSumsD, main = "School D")

StudentSumsE<-colSums(storytellerE[3:8])
StudentSumsE

barplot(StudentSumsE, main = "School E")

```

# Initial Observations and Remarks

Observe:

It’s understood that most of the students are behind even in the breakdown of each school. However, we do notice that school B has more students who completed the course.

This plot has presented the students who have completed the course from school B by section. We found that section 6, 10, and 12 have the better results within school B.

```{r finalANA, include=TRUE}

#looking at the data to determine quality sections

barplot(storytellerB$Completed, names.arg =c(1:12),ylim=c(0,30), main = "Students Completed by Section")
```

This plot has presented the students who are behind the course from school B by section. We found that section 1, 3, and 9 have higher numbers of being behind within school B.

```{r final2, include=TRUE}
rowSums(storytellerB[,3:8])

barplot(storytellerB$`Behind -1-5`+storytellerB$`More Behind -6-10`+storytellerB$`Very Behind -11`, names.arg =c(1:12), ylim=c(0,30), main = "Students Behind by Section")
```

# Results
With exploratory analysis, we have below important findings.
1.	The students' learning abilities and performance are not normally distributed across schools and sections.
2.	By reviewing the boxplot, we understand that the average number of students is higher in the Behind category.
3.	There are not many outliers in students' learning progress. By looking at the log transformation, we can see one or two outliers in completed, middling, behind, and very behind categories.
4.	The distribution of learning progresses is very similar across all five schools.
5.	Over 47% of students are behind, 6% of students are more behind, and 13% of students are very behind. In other words, there are almost 70% of students who are behind in the course.
6.	School B has a greater number of students who completed the course.


# Conclusions

With exploratory data analysis, the researcher can understand how the students are performing in the course. It also gave the researcher an opportunity to explore the data and ask questions such as – which school has students with better performance? By observing the data as a third party, we analyze the data with statistical methods. We found that school B has a higher percentage of students who completed the course. It's worth exploring how school B achieves better performance and if it can be applied to other schools.
